Analysis of bulk void regions

### 2014-04-29 cosmic_voids.tex
Analysis of FA field slides

Once defined the numerical details about both classification schemes, we
shall analyse the dependence on the threshold value $\lambda_{th}$ for 
each one. For this, we shall use the distribution of dark matter halos as 
tracer of the underlying matter field in order to be more consistent with
available observational data. In Figure \ref{fig:halos_fractions} we 
calculate fractions of halos within each one of the defined environments 
based upon the FOF catalogue of the simulation and for an extensive \lth 
range. Then we look for some key feature that could indicated us a 
possible optimal value of the \lth value. One first step forward our quest 
is the behaviour of the V-web scheme compared with the T-web. As was 
previously established by \SRKED{Hoffman et al. (2012)} and as can be seen 
in Figure \ref{fig:halos_fractions}, V-web scheme is significantly more 
sensible to variations of the $\lambda_{th}$ value, since all fractions of
halos for the V-web change significantly in the range $[0,0.4]$, whereas, 
for the T-web scheme, fractions change smoothly throughout all \lth range 
covered. From this, it is then expected that the optimal \lth value for 
the V-web scheme is less than the T-web value.

The more notorious characteristic of Figure \ref{fig:halos_fractions} is
the behaviour of the fraction of halos within sheet regions for both web 
schemes, increasing until a local maximum, and then decreasing. The 
increasing or decreasing rate of the fraction of halos for some region 
could be interpreted as a measure of the degree of non-linearity of such 
region for some specific \lth value. For example, filaments and knots, 
that are the most non-linear regions of the universe, have a negative rate
for all covered \lth range. In the case of voids, the situation is 
completely opposite, where fractions of halos increase in everywhere. If 
we think in terms of the underlying matter field of the cosmic web, \lth 
is just a cutting parameter between high non-linear regions (filament and 
knots) and low non-linear (voids and sheets). Furthermore, if we take into 
account that dark matter halos are much more likely to form in high 
non-linear regions, it is expected the obtained behaviour of fractions of 
halos for voids, filaments and knots as we increase the \lth value. 
However, the behaviour of the fraction of halos in sheets is less clear,
increasing for low \lth values (like voids) and decreasing for higher \lth 
values (like filaments and knots). This indicates us the transitional 
character of sheet regions in the cosmic web. Our proposal here is to 
select as optimal \lth the value where the fraction of sheets reaches a
local maximum, so sheets are completely taken as intermediate transitional 
and neutral zones regarding the degree of non-linearity. According to this,
we find for the T-web scheme an optimal value $\lambda_{opt}^T = 0.36$ and
for the V-web scheme $\lambda_{opt}^V = 0.20$. In figure 
\ref{fig:visual_impression} we show the visual impression of the cosmic 
web along with the density field for different \lth values including the 
optimal values. It can be noticed that the optimal values found reproduce 
well the visual impression.

As we have taken halos as tracers of the cosmic web, we analyse 
distributions of mass and peculiar velocity in order to assign typical 
values to each type of environment. In figure 
\ref{fig:typical_mass_velocity} we calculate both distributions for web 
schemes and using the FOF catalogue of the simulation. Thick lines 
correspond to the median of the distribution and filled regions limited by
dashed lines correspond to quartiles $Q_1$ and $Q_3$, it means, $50\%$ of
all halos are within such regions for every \lth value and for each type 
of environment. We rather use median and quartiles as measure of 
dispersion because there are some very unusual and extreme values that 
makes the usual analysis based upon means and standard deviations less 
reliable.

A first interesting feature of Figure  \ref{fig:typical_mass_velocity} is 
the median mass for each region. In the case of the T-web, although 
dispersions of the distribution of mass for each environment are 
considerably overlapped each other, the median value is very 
well-differentiated among types of environment, indicating that it is 
possible to assign typical values of mass to each region, and being 
consistent with expectations, where low mass halos are typical in voids 
until high mass halos in knots. For the case of the V-web scheme, all
medians and dispersions are completely overlapped, specially for values 
grater than the optimal \lth value, indicating that it is not possible to
assign typical mass ranges to each environment as quantified by this 
scheme. For peculiar velocities, this situation is opposite, where V-web
scheme is much more adequate to assign typical distributions of velocity
to each environment. Although T-web also makes a differentiation in the 
distributions of velocity, this is very slight compared with the V-web 
case. These results can be explained by appealing the physical origin of 
each web scheme. As T-web is based upon the Hessian matrix of the 
potential field, it is expected all quantities related to the potential, 
like density field and distribution of halos mass, are well-differentiated 
among each region, while for the V-web scheme, based upon the shear 
velocity tensor, all dynamical quantities, as the peculiar velocity field 
and the distribution of halos velocity are alike expected to be 
well-differentiated among regions as quantified by this scheme.

Finally, we also calculate typical distributions of the density and 
peculiar velocity fields in each type of environment, obtaining completely
analogous results. Furthermore, we also use a BDM catalogue of the 
simulation, obtaining very similar conclusions.

@@

%-------------------------------------------------------------------------
\subsection{FOF void finder}
\label{subsec:fof_finder}
%-------------------------------------------------------------------------

Following the recent growing interest in studying galaxy formation in 
low-density regions as cosmological tests, we propose an scheme to find 
bulk voids by using the two defined web schemes. Initially, we explore a 
first simple method based on a FOF-like algorithm, where we build an input 
catalogue for the FOF method with the coordinate of the center of every 
cell marked as void according to the web classification scheme adopted, 
setting a linking length to connect diagonal neighbour cells.

Following the work of \SRKED{Forero-Romero et al. 2008}, we also perform a
percolation analysis in order to select the best threshold parameter that
reduces percolation in cells. In Figure \ref{fig:percolation_analysis_FOF}
we show the obtained result of our percolation analysis for both web 
schemes. In both cases, it can be noticed that the volume of the largest 
void region is minimized at $\lambda_{th} = 0.0$, what means the 
percolation is completely reduced for this threshold value. However, 
according to our previous analysis, such value does not reproduce the
visual appearance of the cosmic web, so it is out of consideration.

%*************************************************************************
\section{Statistics of voids and influence over dark matter halos}
\label{sec:statistics}
%*************************************************************************

@@

%.........................................................................
%FIGURE 1: Fraction of halos in each environment
\begin{flushleft}
\begin{figure}
\centering

  \includegraphics[trim = 0mm 0mm 5mm 5mm, clip, keepaspectratio=true,
  width=0.25\textheight]{./figures/halos_fraction_FOF_Tweb.pdf}

  \includegraphics[trim = 0mm 0mm 5mm 5mm, clip, keepaspectratio=true,
  width=0.25\textheight]{./figures/halos_fraction_FOF_Vweb.pdf}

  \captionof{figure}{\small Fractions of halos embedded in each one of 
  the defined environments according to the \lth value. T-web scheme 
  (upper panel) and V-web scheme (lower panel). The optimal parameters 
  found are $\lambda_{opt}^{T}=0.36$ and $\lambda_{opt}^{V}=0.20$.}

  \label{fig:halos_fractions}
  \vspace{0.1 cm}

\end{figure}
\end{flushleft}
%.........................................................................

%.........................................................................
%FIGURE 2: Visual impression of the cosmic web
\begin{flushleft}
\begin{figure*}
\centering

  \includegraphics[trim = 42mm 10mm 37mm 10mm, clip, keepaspectratio=true,
  width=0.75\textheight]{./figures/cosmicweb_visual_Tweb.pdf}
  \includegraphics[trim = 42mm 10mm 37mm 10mm, clip, keepaspectratio=true,
  width=0.75\textheight]{./figures/cosmicweb_visual_Vweb.pdf}

  \captionof{figure}{\small Visual impression of the density field (left 
  panels), and of each classification scheme with the $\lambda_{th}$ values 
  obtained by our criteria (others panels). The color convention for each 
  environment is (white) - void, (light gray) - sheet, (gray) - filament, 
  (black) - knot. For each web scheme, it has been used the previously 
  established optimal threshold as a reference value, so plots are done 
  with the next values $\lambda_{th} = 0.0$, $\lambda_{th} = 
  \lambda_{opt}/2$, $\lambda_{th} = \lambda_{opt}$ and $\lambda_{th} = 
  2\lambda_{opt}$.}

  \label{fig:visual_impression}
  \vspace{0.1 cm}

%.........................................................................
%FIGURE 3: Typical mass and pecular velocity
\begin{flushleft}
\begin{figure*}
\centering

@@

  \captionof{figure}{\small Distribution of masses of dark matter halos
  according the region where they are embedded for both web schemes 
  (upper panels) and of peculiar velocity (lower panels).}

@@

%.........................................................................
%FIGURE 4: Percolation analysis for FOF void finder
\begin{flushleft}
\begin{figure*}
\centering

  \includegraphics[trim = 1mm 10mm 8mm 18mm, clip, keepaspectratio=true,
  width=0.25\textheight]{./figures/voids_regions_percolation_FOF.pdf}

  \captionof{figure}{\small Percolation analysis of bulk void found by using
  FOF void finder algorithm. It is swept an extensive \lth range for both web
  schemes. T-web (blue lines) and  V-web (green lines). Plot of the largest 
  volume (lower panel) and the number of bulk voids found (upper panel).}

  \label{fig:percolation_analysis_FOF}
  \vspace{0.1 cm}

%.........................................................................
%FIGURE 5: Percolation analysis for FOF void finder
\begin{flushleft}
\begin{figure*}
\centering

  \includegraphics[trim = 0mm 0mm 5mm 0mm, clip, keepaspectratio=true,
  width=0.25\textheight]{./figures/voids_regions_volume_FOF_null.pdf}
  \includegraphics[trim = 0mm 0mm 5mm 0mm, clip, keepaspectratio=true,
  width=0.25\textheight]{./figures/voids_regions_volume_FOF_opt.pdf}

  \captionof{figure}{\small Percolation analysis of bulk void found by using
  FOF void finder algorithm. It is swept an extensive \lth range for both web
  schemes. T-web (blue lines) and  V-web (green lines). Plot of the largest 
  volume (lower panel) and the number of bulk voids found (upper panel).}

  \label{fig:volume_distributions}
  \vspace{0.1 cm}

>>>>

Once defined the numerical details about both classification schemes, we
shall analyse the dependence on the threshold value $\lambda_{th}$ for 
each one. For this, we shall use the distribution of dark matter halos as 
tracer of the underlying matter field in order to be more consistent with
available observational data. First, we analyse distributions of mass and 
peculiar velocity in order to assign typical values to each type of 
environment. In the figure \ref{fig:typical_mass_velocity} we calculate 
both distributions for web schemes and using the FOF catalogue of the 
simulation. Thick lines correspond to the median of the distribution and 
filled regions limited by dashed lines correspond to quartiles $Q_1$ and 
$Q_3$, it means, $50\%$ of all halos are within such regions for every 
\lth value and for each type of environment. We rather use median and 
quartiles as measure of dispersion because there are some very unusual and 
extreme values that makes the usual analysis based upon means and standard 
deviations less reliable.

A first interesting feature of the figure  \ref{fig:typical_mass_velocity} 
is the median mass for each region. In the case of the T-web, although 
dispersions of the distribution of mass for each environment are 
considerably overlapped each other, the median value is very 
well-differentiated among types of environment, indicating that it is 
possible to assign typical values of mass to each region, and being 
consistent with expectations, where low mass halos are typical in voids 
until high mass halos in knots. For the case of the V-web scheme, all
medians and dispersions are completely overlapped, specially for values 
grater than the optimal \lth value, indicating that it is not possible to
assign typical mass ranges to each environment as quantified by this 
scheme. For peculiar velocities, this situation is opposite, where V-web
scheme is much more adequate to assign typical distributions of velocity
to each environment. Although T-web also makes a differentiation in the 
distributions of velocity, this is very slight compared with the V-web 
case. These results can be explained by appealing the physical origin of 
each web scheme. As T-web is based upon the Hessian matrix of the 
potential field, it is expected all quantities related to the potential, 
like density field and distribution of halos mass, are well-differentiated 
among each region, while for the V-web scheme, based upon the shear 
velocity tensor, all dynamical quantities, as the peculiar velocity field 
and the distribution of halos velocity are alike expected to be 
well-differentiated among regions as quantified by this scheme.

Finally, we also calculate typical distributions of the density and 
peculiar velocity fields on the volume cells associated to each type of 
environment, obtaining completely analogous results. Furthermore, we use a 
BDM catalogue of the simulation as well, obtaining very similar conclusions.

@@

According to the recent growing interest in studying galaxy formation in 
low-density regions as cosmological tests, classifying void regions is 
becoming an important task in cosmology. Most of those classification 
schemes for voids in cosmological simulations are based upon the density 
field, setting a cut off value below which some region becomes a void 
\SRKED{[references]}. Some more advanced classification schemes are based 
on Voronoi tessellations applied over the tracer particles of the 
simulation in order to compute the density field. Then, through a watershed 
transform, a hierarchy of void regions are found \SRKED{[references, 
ZOBOV algorithm]}.

As has been established \SRKED{[references]}, both schemes presented in
the previous section (V-web and T-web) for classifying the cosmic web 
present many advantages compared with classification schemes based
completely upon the density field, e.g. a more robust description of the
dynamic a kinematic of the cosmic web, a more reliable quantification of 
the visual impression, among others. With the aim of exploiting all of 
these advantages, we propose here a novel approach to classify voids in
cosmological simulations based entirely on the web schemes.

The original version of the T-web scheme \SRKED{[reference, Hahn]} was not
successful at reproducing the visual impression of the cosmic web, 
however, with the introduction of a threshold parameter \SRKED{[reference, 
Forero-Romero]}, this scheme, and even the V-web \SRKED{[reference, 
Hoffman]}, improved enormously. As this free parameter controls the visual 
impression provided by each scheme, phenomenons like percolation depends 
on it as well. Percolation is one of the key features of the structure of 
void regions, indicating how voids are merged among them, and how they 
permeate all the cosmic web.

In order to overcome the percolation of voids in our classification scheme,
we introduce the fractional anisotropy as defined in \SRKED{[reference, 
Libeskind]}.

%.........................................................................
%Fractional anisotropy
\eq{fractional_anisotropy}
{ FA = \frac{1}{\sqrt{3}}\sqrt{ \frac{ (\lambda_1 - \lambda_3)^2 + 
(\lambda_2 - \lambda_3)^2 + (\lambda_1 - \lambda_2)^2}{ \lambda_1^2 + 
\lambda_2^2 + \lambda_3^2} } }
%.........................................................................
where the eigenvalues are taken from any of the two web schemes. This 
index, such as it is defined, allows quantifying the local anisotropy 
degree of the cosmological environment, where $FA=0$ corresponds to highly
isotropic regions and $FA=1$ anisotropic ones.

In the figure \ref{fig:FA_field} we calculate the FA field over the 
simulation for both web schemes. The first interesting feature of this
figure is the degeneration presented for knots and central regions of 
voids, where both of them exhibit low to middle values of the FA, 
indicating a high isotropy regarding the physical property quantified by 
each web scheme, i.e. the density field for the T-web and the peculiar 
velocity for the V-web. For the T-web, the FA field around knots presents 
a very narrow distribution, whereas for the V-web the same distribution is
more spread. This can be explained appealing to the low dispersion of the 
density field with respect to the peculiar velocity in highly non-linear 
regions like knots. For more linear regions like voids, the behaviour of 
the FA field is quite similar between both schemes, what is consistent 
with the equivalence of the T-web and the V-web in the linear regime 
\SRKED{[reference, Hoffman]}.

According to the classification scheme adopted for the cosmological 
environment, voids are regions where 
$\lambda_3\leq\lambda_2\leq\lambda_1\leq\lambda_{th}$, this implies that
the boundaries of void regions are controlled completely by the $\lambda_1$
eigenvalue of the web scheme and the threshold value. Therefore, as we 
increase the threshold value $\lambda_th$, all voids grow up progressively 
through contours of $\lambda_3$ field until certain critical value where 
voids are very large so the visual impression is no longer reliable.

%*************************************************************************
\section{Properties of voids}
\label{sec:properties}
%*************************************************************************

Once defined the proper scheme to classify bulk voids in the simulation,
we proceed to analyse their physical properties, like the inertia values,
the density and peculiar velocities profiles as calculated over the grid 
and profiles of number of halos.

%-------------------------------------------------------------------------
\subsection{Shape of voids}
\label{subsec:shape_voids}
%-------------------------------------------------------------------------

Quantifying the shape of voids is gaining importance due to cosmological 
tests such as the Alcock-Paczynski test \SRKED{[Sutter, et.al (2012)]}, so 
we compute here the reduced inertia tensor through the next expression in 
order to determine shape distributions of bulk voids.

%.........................................................................
%Reduced inertia tensor
\eq{ReducedIntertia}
{ \tau_{ij} = \sum_l \frac{ x_{l,i}x_{l,j}  }{R_l^2} }
%.........................................................................
where $l$ is an index associated to each cell of the current region, 
$i$ and $j$ indexes run over each spatial direction and finally 
$R_l$ is defined as $R_l^2 = x_{l,1}^2 + x_{l,2}^2 + x_{l,3}^2$. All 
positions are measured from the respective geometric center of each void.

The eigenvalues of the reduced inertia tensor, i.e. the principal moments
of inertia, are used to quantify the shape of each bulk void. They are 
denoted as $\tau_1$, $\tau_2$ and $\tau_3$ such that $\tau_1 \leq \tau_2
\leq \tau_3$. In Figure \ref{fig:distro_inertia} we show the computed
distributions for $\tau_1/\tau_2$ and $\tau_2/\tau_3$ for voids larger 
than 8 cells in order to avoid statistic fluctuations due to small regions.
We rather calculate histograms for these ratio quantities instead of each 
single value in order to avoid using an arbitrary normalization. For both 
schemes, it can be noticed that the shape distribution is completely 
spread out, thereby indicating a non-preferred geometry of void regions, 
which is in agreement with the well established high anisotropy of matter 
flows associated to this type of region.

For a better quantification, we also perform a classification of the shape 
of voids by setting a threshold in the analysed ratio quantities. An 
anisotropic or tri-axial shape correspond to voids where $\tau_1/\tau_2 < 
0.7$ and $\tau_2/\tau_3 < 0.7$, where there is not any symmetry among the
principal directions. We find about $57.2\% \sim 61.0\%$ of total voids 
consistent with this shape, for the T-web and V-web respectively. A 
pancake or quasi-oblate shape is associated to voids where $\tau_1/\tau_2 
< 0.7$ and $\tau_2/\tau_3 > 0.7$. We found $13.1\% \sim 17.9\%$ of 
consistent voids. Filamentary or quasi-prolate voids satisfy $\tau_1/\tau_2 
> 0.7$ and $\tau_2/\tau_3 < 0.7$, with $25.4\% \sim 18.1\%$ of all voids.
Finally, isotropic or quasi-spheric voids are found when $\tau_1/\tau_2 
> 0.7$ and $\tau_2/\tau_3 > 0.7$, with $4.2\% \sim 3.1\%$ of total voids 
compatible with this shape. The threshold value of $0.7$ adopted here for
the ratios of the moments of inertia is just for illustrative purposes, 
where such distinction is rather fuzzy and continuous. However, the 
previous analysis allows us to conclude that voids are quite asymmetric 
structures.

%-------------------------------------------------------------------------
\subsection{Density profile of voids}
\label{subsec:density_voids}
%-------------------------------------------------------------------------

Describing the density profiles of voids is quite important in order to 
compare and match simulation with observational surveys, allowing possible
constrains for different cosmology models \SRKED{[Hamaous, et.al 2014]}. 
Here, and taking into account the previous results, we rather use an 
ellipsoidal approximation to describe and fit the shape of bulk voids, so 
we use the next ellipsoidal radial coordinate to describe density profiles.

%.........................................................................
%Ellipsoidal radial coordinate
\eq{radial_coordinate}
{
r^2 = \frac{x^2}{\tau_1^2} + \frac{y^2}{\tau_2^2} + \frac{z^2}{\tau_3^2},
\ \ \ \ 0\leq r \leq 1
}
%.........................................................................
where we take the principal moments of inertia $\{\tau_i \}$ as the 
lengths of the principal axes of the ellipsoid and each one of the 
cartesian coordinates as measured in the rotated frame of each void.

We use the same analytic density profile that \SRKED{[Hamaous, et.al 2014]} 
to fit the numerical density profiles of our voids.

%.........................................................................
%Density profile
\eq{density_profile}
{
\delta_v(r) = \delta_c\frac{1-(r/r_s)^\alpha}{1+(r/r_v)^\beta}
}
%.........................................................................

@@

%.........................................................................
%FIGURE 1: Typical mass and pecular velocity
\begin{flushleft}
\begin{figure*}
\centering

@@

  \captionof{figure}{\small Distribution of masses of dark matter halos
  according the region where they are embedded for both web schemes 
  (upper panels) and of peculiar velocity (lower panels). It can be 
  noticed that the T-web scheme selects different mass ranges according
  to the environment, while the V-web scheme is better selecting ranges
  of peculiar velocity of the dark matter halos. This can be understood
  taking into account the physical origin of each web scheme.}

@@

%.........................................................................
%FIGURE 2: Typical mass and pecular velocity
\begin{flushleft}
\begin{figure*}
\centering

  \includegraphics[trim = 25mm 200mm 18mm 15mm, clip, keepaspectratio=true,
  width=0.5\textheight]{./figures/cosmicweb_FA_Tweb(null).pdf}

  \includegraphics[trim = 25mm 200mm 18mm 15mm, clip, keepaspectratio=true,
  width=0.5\textheight]{./figures/cosmicweb_FA_Vweb(null).pdf}

  \captionof{figure}{\small In left panels is shown the visual impression 
  of the cosmic web for each web scheme (T-web, upper panels. V-web, lower 
  panels) obtained for $\lambda_{th}=0.0$. It can be seen each one of the 
  defined type of environments, where voids corresponds to white zones, 
  sheets to gray, filaments to dark gray and finally knots to black regions. 
  In the right panels is shown the fractional anisotropy field for the same 
  slide of the simulation and for each web schemes, where black regions 
  correspond to FA$=1$ and white regions to FA$=0$. It can be noticed the 
  degeneration of low values of FA for knots and central regions of voids, 
  while high values of FA (FA $\lesssim 1$) are consistent with filaments 
  and highly planar sheets.}

  \label{fig:FA_field}
  \vspace{0.1 cm}
